R Notebook

### 2022-03-31 Code/Simulation.Rmd
Added GEV testing function, comments in code, and correct naive functions

```{r}
# Load packages
library(xtable)
library(dgof)
library(tidyverse)
library(KSgeneral)
library(copula)
source('tspb.R')
```

@@

  # Parametric bootstrap to approximate the null distribution
  n <- length(x)
  stat.b <- double(B)
  for (i in 1:B) {
    x.b <- rnorm(n, mu, sigma)
    mu.b <- mean(x.b)
    sigma.b <- sd(x.b)
    stat.b[i] <- ks.test(x.b, "pnorm", mu.b, sigma.b)$statistic
  }
  p.value <- (sum(stat.b >= stat) + 0.5) / (B + 1)

@@

test_babu <- function(x, B = 1000) {
  p.value <- ks.np.babu(x, B = 1000)$p.value
  counter <<- counter + 1
  cat(counter, "")
  list(p.value = p.value)
}

@@

```{r}
test_copula <- function(x, dist, B) {
    ## example assuming normal distribution; normal copula
    ## get observed stat
    n <- length(x)
    mu <- mean(x)
    sigma <- sd(x)
    stat  <- ks.test(x, dist, mean = mu, sd = sigma)$statistic
    ## get lag-1 sample auto-spearman rho
    rho  <-  cor(x[-1], x[-n], method = "spearman")
    r <- iRho(normalCopula(), rho)
    np <- normalCopula(r)
    ## parametric bootstrap to get an empirical distribution of stat
    stat.b <- double(n)
    ## set up containers
    for (i in 1:B) {
        ## get one bootstrap sample
        ## 1. the marginal distribution is the fitted dist
        ## 2. there is temporal dependence
        u <- runif(n) 
        for (j in 2:n) {
            ## could test other copulas later
            u[j]  <- cCopula(c(u[j-1], u[j]), np, indices = 2, inverse = TRUE)
        }
        ## get bootstrap sample
        x.b <- qnorm(u, mean = mu, sd = sigma)
        ## fit it
        mu.b <- mean(x.b)
        sigma.b <- sd(x.b)
        stat.b[i] <- ks.test(x.b, dist, mean = mu.b, sd = sigma.b)$statistic
    }

    p.value <- (sum(stat.b >= stat) + 0.5) / (B + 1)

@@

    return(p.value)
    # return empirical p-value    
}

@@

# Generate results
n <- 100
ar <- 0.5
counter <- 0
nrep <- 1000
B <- 1000
pvals_copula <- replicate(nrep, test_copula(arima.sim(list(ar = ar), rand.gen=rnorm, sd = sqrt(1 - ar^2), n = n), "pnorm", B = B))
pvals_naive <- replicate(nrep, test_naive_copula(arima.sim(list(ar = ar), rand.gen=rnorm, sd = sqrt(1 - ar^2), n = n))$p.value)

@@

# Generate results
n <- 100
counter <- 0
nrep <- 1000
B <- 100
b1 <- 0.5
b2 <- 0.3
pvals_copula <- replicate(nrep, test_copula(arima.sim(list(order = c(2, 0, 0), 
                                                           ar = c(b1, b2)), 
                                                          rand.gen=rnorm, 
                                                          sd = sqrt(((1+b2)*(1-b2-b1)*(1-b2+b1))/(1-b2)), 
                                                          n = n), "pnorm", B = B))
pvals_naive <- replicate(nrep, test_naive_copula(arima.sim(list(order = c(2, 0, 0), ar = c(b1, b2)), rand.gen=rnorm, sd = sqrt(((1+b2)*(1-b2-b1)*(1-b2+b1))/(1-b2)), n = n))$p.value)

@@

# Save the histogram figure
#ggsave(filename = "hist_copula_ar2.pdf", plot = hist_copula_ar2, path = "../Manuscript", height = 2.1, width = 6)
```

```{r}
b1 <- 0.5
b2 <- 0.3
n <- 100
x <- arima.sim(list(order = c(2, 0, 0), ar = c(b1, b2)), rand.gen=rnorm, sd = sqrt(((1+b2)*(1-b2-b1)*(1-b2+b1))/(1-b2)), n = n)
sd(x)
var(x)

ma <- 0.5
x <- arima.sim(list(ma = 0.5), rand.gen = rnorm, sd = sqrt(1/(1+ma^2)), n = n)
sd(x)
var(x)

ar <- 0.5
x <- arima.sim(list(ar = ar, ma = ma), rand.gen = rnorm, sd = sqrt((1-(ar^2))/(1+2*ar*ma+ma^2)), n = n)
sd(x)
var(x)
```

@@

# Generate results
n <- 100
counter <- 0
nrep <- 1000
B <- 100
ma <- 0.5
pvals_copula <- replicate(nrep, test_copula(arima.sim(list(ma = 0.5), rand.gen = rnorm, sd = sqrt(1/(1+ma^2)), n = n), "pnorm", B = B))
pvals_naive <- replicate(nrep, test_naive_copula(arima.sim(list(ma = 0.5), rand.gen = rnorm, sd = sqrt(1/(1+ma^2)), n = n))$p.value)

@@

# Generate results
n <- 100
counter <- 0
nrep <- 1000
B <- 100
ar <- 0.5
ma <- 0.3
pvals_copula <- replicate(nrep, test_copula(arima.sim(list(ar = ar, 
                                                           ma = ma), 
                                                      rand.gen=rnorm, 
                                                      sd = sqrt((1-(ar^2))/(1+2*ar*ma+ma^2)), 
                                                      n = n),
                                            "pnorm", B = B))
pvals_naive <- replicate(nrep, test_naive_copula(arima.sim(list(ar = ar, 
                                                                ma = ma), 
                                                      rand.gen=rnorm, 
                                                      sd = sqrt((1-(ar^2))/(1+2*ar*ma+ma^2)), 
                                                      n = n))$p.value)

@@

```{r}
library(fitdistrplus)
test_gamma <- function(x, dist, B = 100, wcopula = normalCopula) {

    #fit <- fitdist(x, "gamma", method = "mle", lower = c(0, 0), start = list(scale = 1, shape = 1))
    #shape <- fit$estimate[1]
    #scale <- fit$estimate[2]

@@

    ## example assuming normal distribution; normal copula
    ## get observed stat
    n <- length(x)

    ## get lag-1 sample auto-spearman rho
    rho  <-  cor(x[-1], x[-n], method = "spearman")
    r <- iRho(wcopula(), rho)
    wcop <- wcopula(r)
    ## parametric bootstrap to get an empirical distribution of stat
    stat.b <- double(n)
    ## set up containers
    for (i in 1:B) {
        ## get one bootstrap sample
        ## 1. the marginal distribution is the fitted dist
        ## 2. there is temporal dependence
        u <- runif(n)
        for (j in 2:n) {
            ## could test other copulas later
            u[j]  <- cCopula(c(u[j-1], u[j]), wcop, indices = 2, inverse = TRUE)
        }
        ## get bootstrap sample
        ## get fitted parameters for the bootstrap sample

        x.b <- qgamma(u, shape = shape, scale = scale)

@@

        #fit.b <- fitdist(x.b, "gamma", method = "mle", lower = c(0, 0), start = list(scale = 1, shape = 1))
        #shape.b <- fit.b$estimate[1]
        #scale.b <- fit.b$estimate[2]

        stat.b[i] <- ks.test(x.b, dist, shape = shape.b, scale = scale.b)$statistic
    }
    ## return empirical p-value    
    p.value <- (sum(stat.b >= stat) + 0.5) / (B + 1)

# Naive KS-test
test_naive_copula <- function(x) {
  p.value <- ks.test(x, "pgamma", shape = 3, scale = 1)$p.value
  list(p.value = p.value)
}

@@

# Generate results
n <- 100
rho <- 0.5
counter <- 0
nrep <- 1000
B <- 1000
pvals_copula <- replicate(nrep, test_gamma(genData(n, qgamma, shape = 3, scale = 1, copula = normalCopula, rho = rho), "pgamma", B = B))
pvals_naive <- replicate(nrep, test_naive_copula(genData(n, qgamma, shape = 3, scale = 1, copula = normalCopula, rho = rho))$p.value)

@@

```{r}
rho <- 0.5
x <- genData(100, qgamma, shape = 3, rate = 2, copula = normalCopula, rho = rho)
ks.test(x, "pgamma", shape = 3, rate = 2)
```
}}

>>>>

```{r}
# Load packages
library(xtable)
library(dgof)
library(tidyverse)
library(KSgeneral)
library(copula)
library(evd)
source('tspb.R')
```

@@

  # Parametric bootstrap to approximate the null distribution
  n <- length(x)
  stat.b <- double(B)
  for (i in 1:B) {
    x.b <- rnorm(n, mu, sigma)
    mu.b <- mean(x.b)
    sigma.b <- sd(x.b)
    stat.b[i] <- ks.test(x.b, "pnorm", mu.b, sigma.b)$statistic
  }

  # Return empirical p-value
  p.value <- (sum(stat.b >= stat) + 0.5) / (B + 1)

@@

# Test non-parametric bootstrap as a solution to fitted parameters
test_babu <- function(x, B = 1000) {
  p.value <- ks.np.babu(x, B = 1000)$p.value
  counter <<- counter + 1
  cat(counter, "")
  list(p.value = p.value)
}

@@

```{r}
test_copula <- function(x, dist, B) {

    # Get observed statistic
    n <- length(x)
    mu <- mean(x)
    sigma <- sd(x)
    stat  <- ks.test(x, dist, mean = mu, sd = sigma)$statistic

    # Get lag-1 sample auto-spearman rho
    rho  <-  cor(x[-1], x[-n], method = "spearman")
    r <- iRho(normalCopula(), rho)
    np <- normalCopula(r)

    # Perform parametric bootstrap to get an empirical distribution of stat
    stat.b <- double(n)

    for (i in 1:B) {
        # 1. The marginal distribution is the fitted dist
        # 2. There is temporal dependence
        u <- runif(n) 
        for (j in 2:n) {
            u[j]  <- cCopula(c(u[j-1], u[j]), np, indices = 2, inverse = TRUE)
        }

        # Transform bootstrap sample and fit the parameters
        x.b <- qnorm(u, mean = mu, sd = sigma)
        mu.b <- mean(x.b)
        sigma.b <- sd(x.b)

        # Gather bootstrap sample's test statistic
        stat.b[i] <- ks.test(x.b, dist, mean = mu.b, sd = sigma.b)$statistic
    }

    # Return empirical p-value
    p.value <- (sum(stat.b >= stat) + 0.5) / (B + 1)

@@

# Super naive KS-test with fitted parameters
test_super_naive <- function(x) {
  p.value <- ks.test(x, "pnorm", mean = mean(x), sd = sd(x))$p.value
  list(p.value = p.value)
}

@@

# Generate results
n <- 100
ar <- 0.5
counter <- 0
nrep <- 1000
B <- 100
pvals_copula <- replicate(nrep, test_copula(arima.sim(list(ar = ar), rand.gen=rnorm, sd = sqrt(1 - ar^2), n = n), "pnorm", B = B))
counter <- 0
pvals_parametric <- replicate(nrep, test_parametric(arima.sim(list(ar = ar), rand.gen=rnorm, sd = sqrt(1 - ar^2), n = n))$p.value)
pvals_naive <- replicate(nrep, test_super_naive(arima.sim(list(ar = ar), rand.gen=rnorm, sd = sqrt(1 - ar^2), n = n))$p.value)

# Create a master data_frame with all the results
histogram <- c(rep("Copula", nrep), rep("Naive Parametric", nrep), rep("Super Naive", nrep))
p_values <- c(pvals_copula, pvals_parametric, pvals_naive)
results <- data_frame(p_values, histogram)

@@

# Set variables
n <- 100
counter <- 0
nrep <- 1000
B <- 100
b1 <- 0.5
b2 <- 0.3

# P-vals from using bootstrap with copulas
pvals_copula <- replicate(nrep, test_copula(arima.sim(list(order = c(2, 0, 0), 
                                                           ar = c(b1, b2)),
                                                      rand.gen=rnorm,
                                                      sd = sqrt(((1+b2)*(1-b2-b1)*(1-b2+b1))/(1-b2)),
                                                      n = n), 
                                            "pnorm", 
                                            B = B))
counter <- 0

# P-vals from using parametric bootstrap
pvals_parametric <- replicate(nrep, test_parametric(arima.sim(list(order = c(2, 0, 0), 
                                                                   ar = c(b1, b2)), 
                                                              rand.gen=rnorm, 
                                                              sd = sqrt(((1+b2)*(1-b2-b1)*(1-b2+b1))/(1-b2)), 
                                                              n = n))$p.value)

# P-vals from naively using fitted parameters
pvals_naive <- replicate(nrep, test_super_naive(arima.sim(list(order = c(2, 0, 0), 
                                                               ar = c(b1, b2)), 
                                                          rand.gen=rnorm, 
                                                          sd = sqrt(((1+b2)*(1-b2-b1)*(1-b2+b1))/(1-b2)), 
                                                          n = n))$p.value)

# Create a master data_frame with all the results
histogram <- c(rep("Copula", nrep), rep("Naive Parametric", nrep), rep("Super Naive", nrep))
p_values <- c(pvals_copula, pvals_parametric, pvals_naive)
results <- data_frame(p_values, histogram)

@@

# Save the histogram figure
ggsave(filename = "hist_copula_ar2.pdf", plot = hist_copula_ar2, path = "../Manuscript", height = 2.1, width = 6)
```

@@

# Set variables
n <- 100
counter <- 0
nrep <- 1000
B <- 100
ma <- 0.5

# P-vals from using bootstrap with copulas
pvals_copula <- replicate(nrep, test_copula(arima.sim(list(ma = 0.5), rand.gen = rnorm, sd = sqrt(1/(1+ma^2)), n = n), "pnorm", B = B))
counter <- 0

# P-vals from using parametric bootstrap
pvals_parametric <- replicate(nrep, test_parametric(arima.sim(list(ma = 0.5), rand.gen = rnorm, sd = sqrt(1/(1+ma^2)), n = n))$p.value)

# P-vals from naively using fitted parameters
pvals_naive <- replicate(nrep, test_super_naive(arima.sim(list(ma = 0.5), rand.gen = rnorm, sd = sqrt(1/(1+ma^2)), n = n))$p.value)

# Create a master data_frame with all the results
histogram <- c(rep("Copula", nrep), rep("Naive Parametric", nrep), rep("Super Naive", nrep))
p_values <- c(pvals_copula, pvals_parametric, pvals_naive)
results <- data_frame(p_values, histogram)

@@

# Set variables
n <- 100
counter <- 0
nrep <- 1000
B <- 100
ar <- 0.5
ma <- 0.3

# P-vals from using bootstrap with copulas
pvals_copula <- replicate(nrep, test_copula(arima.sim(list(ar = ar, 
                                                           ma = ma), 
                                                      rand.gen=rnorm, 
                                                      sd = sqrt((1-(ar^2))/(1+2*ar*ma+ma^2)), 
                                                      n = n),
                                            "pnorm", B = B))
counter <- 0

# P-vals from using parametric bootstrap
pvals_parametric <- replicate(nrep, test_parametric(arima.sim(list(ar = ar, 
                                                              ma = ma), 
                                                      rand.gen=rnorm, 
                                                      sd = sqrt((1-(ar^2))/(1+2*ar*ma+ma^2)), 
                                                      n = n))$p.value)

# P-vals from naively using fitted parameters
pvals_naive <- replicate(nrep, test_super_naive(arima.sim(list(ar = ar, 
                                                              ma = ma), 
                                                      rand.gen=rnorm, 
                                                      sd = sqrt((1-(ar^2))/(1+2*ar*ma+ma^2)), 
                                                      n = n))$p.value)

# Create a master data_frame with all the results
histogram <- c(rep("Copula", nrep), rep("Naive Parametric", nrep), rep("Super Naive", nrep))
p_values <- c(pvals_copula, pvals_parametric, pvals_naive)
results <- data_frame(p_values, histogram)

@@

```{r}
# Testing bootstrap with copulas for gamma distribution
test_gamma <- function(x, dist, B = 100, wcopula = normalCopula) {

    # Get observed stat
    n <- length(x)
    m <- mean(x)
    v <- var(x)
    scale <- v/m
    shape <- m*m/v

@@

    # Get lag-1 sample auto-spearman rho
    rho  <-  cor(x[-1], x[-n], method = "spearman")
    r <- iRho(wcopula(), rho)
    wcop <- wcopula(r)

    # Parametric bootstrap to get an empirical distribution of stat
    stat.b <- double(n)

    for (i in 1:B) {
        # 1. The marginal distribution is the fitted dist
        # 2. There is temporal dependence
        u <- runif(n)
        for (j in 2:n) {
            u[j]  <- cCopula(c(u[j-1], u[j]), wcop, indices = 2, inverse = TRUE)
        }

        # Transform the bootstrap sample and fit the parameters
        x.b <- qgamma(u, shape = shape, scale = scale)

        m.b <- mean(x.b)
        v.b <- var(x.b)
        scale.b <- v.b/m.b
        shape.b <- m.b*m.b/v.b

        # Calculate new test statistic
        stat.b[i] <- ks.test(x.b, dist, shape = shape.b, scale = scale.b)$statistic
    }

    # Return empirical p-value    
    p.value <- (sum(stat.b >= stat) + 0.5) / (B + 1)

# Parametric bootstrap to solve fitted parameters for gamma distribution
test_parametric_gamma <- function(x, B = 1000) {

  # Summarize parameters
  m <- mean(x)
  v <- var(x)
  scale <- v/m
  shape <- m*m/v

  # Store KS statistic from using fitted parameters
  stat <- ks.test(x, "pgamma", shape = shape, scale = scale)$statistic

  # Parametric bootstrap to approximate the null distribution
  n <- length(x)
  stat.b <- double(B)
  for (i in 1:B) {
    x.b <- rgamma(n, shape = shape, scale = scale)

@@

    stat.b[i] <- ks.test(x.b, "pgamma", shape = shape.b, scale = scale.b)$statistic
  }

  # Return empirical p-value
  p.value <- (sum(stat.b >= stat) + 0.5) / (B + 1)

# Naive function that incorrectly uses fitted parameters for gamma distribution
test_super_naive_gamma <- function(x) {

  p.value <- ks.test(x, "pgamma", shape = shape, scale = scale)$p.value
  list(p.value = p.value)
}

@@

# Set variables
n <- 100
rho <- 0.7
counter <- 0
nrep <- 1000
B <- 100

# P-vals from using bootstrap with copulas
pvals_copula <- replicate(nrep, test_gamma(genData(n, qgamma, shape = 3, scale = 1, copula = normalCopula, rho = rho), "pgamma", B = B))
counter <- 0

# P-vals from using parametric bootstrap
pvals_parametric <- replicate(nrep, test_parametric_gamma(genData(n, qgamma, shape = 3, scale = 1, copula = normalCopula, rho = rho))$p.value)

# P-vals from naively using fitted parameters
pvals_naive <- replicate(nrep, test_super_naive_gamma(genData(n, qgamma, shape = 3, scale = 1, copula = normalCopula, rho = rho))$p.value)

# Create a master data_frame with all the results
histogram <- c(rep("Copula", nrep), rep("Naive Parametric", nrep), rep("Super Naive", nrep))
p_values <- c(pvals_copula, pvals_parametric, pvals_naive)
results <- data_frame(p_values, histogram)

@@

```{r}
# Testing bootstrap with copulas for GEV distribution
test_gev <- function(x, B = 100, wcopula = normalCopula) {

    n <- length(x)

    # Get observed stat
    fit <- fgev(x, std.err = FALSE)
    loc <- fit$estimate[1]
    scale <- fit$estimate[2]
    shape <- fit$estimate[3]

    stat  <- ks.test(x, 'pgev', loc = loc, scale = scale, shape = shape)$statistic

    # Get lag-1 sample auto-spearman rho
    rho  <-  cor(x[-1], x[-n], method = "spearman")
    r <- iRho(wcopula(), rho)
    wcop <- wcopula(r)

    # Parametric bootstrap to get an empirical distribution of stat
    stat.b <- double(n)

    for (i in 1:B) {
        # 1. The marginal distribution is the fitted dist
        # 2. There is temporal dependence
        u <- runif(n)
        for (j in 2:n) {
            u[j]  <- cCopula(c(u[j-1], u[j]), wcop, indices = 2, inverse = TRUE)
        }

        # Transform the bootstrap sample and fit the parameters
        x.b <- qgev(u, loc = loc, scale = scale, shape = shape)

        fit.b <- fgev(x.b, std.err = FALSE)
        loc.b <- fit.b$estimate[1]
        scale.b <- fit.b$estimate[2]
        shape.b <- fit.b$estimate[3]

        # Calculate bootstrap test statistic
        stat.b[i] <- ks.test(x.b, 'pgev', loc = loc.b, scale = scale.b, shape = shape.b)$statistic
    }

    # Return empirical p-value    
    p.value <- (sum(stat.b >= stat) + 0.5) / (B + 1)

# Parametric bootstrap to solve fitted parameters for GEV distribution
test_parametric_gev <- function(x, B = 1000) {

  # Summarize parameters
  fit <- fgev(x)
  loc <- fit$estimate[1]
  scale <- fit$estimate[2]
  shape <- fit$estimate[3]

  # Store KS statistic from using fitted parameters
  stat <- ks.test(x, "pgev", loc = loc, scale = scale, shape = shape)$statistic

  # Parametric bootstrap to approximate the null distribution
  n <- length(x)
  stat.b <- double(B)
  for (i in 1:B) {
    x.b <- rgev(n, loc = loc, scale = scale, shape = shape)

    fit.b <- fgev(x.b)
    loc.b <- fit.b$estimate[1]
    scale.b <- fit.b$estimate[2]
    shape.b <- fit.b$estimate[3]

    stat.b[i] <- ks.test(x.b, "pgev", shape = shape.b, scale = scale.b)$statistic
  }

  # Return empirical p-value
  p.value <- (sum(stat.b >= stat) + 0.5) / (B + 1)

# Naive function that incorrectly uses fitted parameters for GEV distribution
test_super_naive_gev <- function(x) {

  fit <- fgev(x)
  loc <- fit$estimate[1]
  scale <- fit$estimate[2]
  shape <- fit$estimate[3]

  p.value <- ks.test(x, "pgev", loc = loc, scale = scale, shape = shape)$p.value
  list(p.value = p.value)
}

# Set variables
n <- 100
rho <- 0.5
counter <- 0
nrep <- 1000
B <- 100

# P-vals from using bootstrap with copulas
pvals_copula <- suppressWarnings(replicate(nrep, test_gev(genData(n, qgev, loc = 0, shape = 1, scale = 1, copula = normalCopula, rho = rho), B = B)))
counter <- 0

# P-vals from using parametric bootstrap
pvals_parametric <- suppressWarnings(replicate(nrep, test_parametric_gamma(genData(n, qgev, shape = 1, scale = 1, copula = normalCopula, rho = rho))$p.value))

# P-vals from naively using fitted parameters
pvals_naive <- suppressWarnings(replicate(nrep, test_super_naive_gamma(genData(n, qgev, shape = 1, scale = 1, copula = normalCopula, rho = rho))$p.value))

# Create a master data_frame with all the results
histogram <- c(rep("Copula", nrep), rep("Naive Parametric", nrep), rep("Super Naive", nrep))
p_values <- c(pvals_copula, pvals_parametric, pvals_naive)
results <- data_frame(p_values, histogram)

# Generate histograms plot
hist_copula_gev <- ggplot(results) + 
                   geom_histogram(aes(x = p_values, fill = histogram), boundary = 0, binwidth = 1/20) + 
                   scale_x_continuous(name = "p-value", breaks=c(0, 1)) + 
                   theme(legend.position = "none") +
                   facet_wrap(~histogram, scales = "free_y")

hist_copula_gev

# Save the histogram figure
ggsave(filename = "hist_copula_gev.pdf", plot = hist_copula_gev, path = "../Manuscript", height = 2.1, width = 6)
```
